### 2014-07-17 supp-power-table.tex

\begin{sidewaystable}[ht]
\begin{center}
\caption{Power of difference in $STG(t)$ tests for $t=3$ years \label{Power}}
\begin{tabular}{rrccccc}
  \hline
 & $R_{s_{1,j}w}$ & $STG(3)=0.1$  & $STG(3)=0.24$  & $STG(3)=0.20$& $STG(3)^{*}=0.28$  & $STG(3)^{*}=0.31$\\
  \hline
$STG(3)=0.1$ &0.8 & 5.5 & 17.0 & 16.8 & 28.2 & 71.0 \\ 
   &0.5 &5.8 & 11.9 & 14.2 & 45.6 & 73.7 \\ 
$STG(3)=0.24$ &0.8&  & 5.1 & 7.9 & 7.5 & 13.7 \\ 
   &0.5 &  & 5.5 & 8.9 & 7.4 & 13.3 \\ 
$STG(3)=0.20$&0.8 &  &  & 5.3 & 52.5 & 96.4 \\ 
   &0.5&  &  & 6.3 & 45.5 & 93.9 \\ 
$STG(3)^{*}=0.28$&(0.8, 0.8) &  &  &  & 5.4 & 7.7 \\ 
   &(0.5, 0.5) &  &  &  & 5.7 & 8.9 \\ 
$STG(3)^{*}=0.31$&(0.8, 0.8)&   &  &  &  & 5.1 \\ 
   &(0.5, 0.5)&  &  &  &  & 4.2 \\ 
   \hline
\multicolumn{7}{p{9in}}{Wald test for comparing two candidate surrogates by testing $STG_1(3)-STG_2(3)=0$. *Multivariate candidate PS, **For multivariate surrogate candidates $\rho_{s_{1,j}w}$ stands for $(R_{s_{1,1}w},R_{s_{1,2}w})$.}
\end{tabular}
\end{center}
\end{sidewaystable}